Over-relaxed Sinkhorn's Algorithm for Regularized Optimal Transport

### 2017-10-18 paper.tex
Algo

{\color{red} 
	TODO Construire les bonnes fonctions $\theta_k(\gamma)$.}
To construct the functions $\theta_k$ needed by the theorem, we need the following lemma.

@@

In practice, we observed that the algorithm introduced in Theorem \ref{thm:algo} can be simplified and it is sufficient to check $F(P^\omega_{\Ccal_2}(P^\omega_{\Ccal_1}(\gamma^l)))-F(\gamma^l)>0$ for performing the over-relaxed iterations with parameter $\omega$.
The cost of this test is linear with respect to the data dimension (i.e. $o(n_1+n_2)$).

\renewcommand{\algorithmiccomment}[1]{\hfill\bgroup(#1)\egroup}
\begin{algorithm}
\caption{Over-relaxed Sinkhorn algorithm}
\label{SOR}
\begin{algorithmic}
\REQUIRE $\mu^1\in \IR^{n_1}$, $\mu^2\in \IR^{n_2}$, $c\in \IR^{n_1\times n_2}_+$
\STATE Set $a=\mathbf{1}_{n_1}$, $b=\mathbf{1}_{n_2}$, $\gamma^0=e^{-c/\epsilon}$, $\omega\in[1;2[$ and $\eta>0$
\WHILE {$||a\otimes \gamma^0b-  \mu_1||>\eta$}
\STATE $\tilde a=\mu_1\oslash (\gamma^0 b)$,  $a_\omega=a^{1-\omega}\otimes \tilde a^\omega$
\STATE $\tilde b=\mu_2\oslash (^t\gamma^0  a_\omega)$
\IF{ $\langle \mu^1,\varphi_\omega(a\oslash\tilde a) \rangle +\langle \mu^2,\varphi_\omega(b\oslash\tilde b)\rangle>0$}
\STATE  $a=a_\omega$, $b=b^{1-\omega}\otimes \tilde b^\omega$\COMMENT{Over-relaxed Sinkhorn iteration}
\ELSE
\STATE $a=\tilde a$, $b=\mu_2\oslash (^t\gamma^0 a)$  \COMMENT{Sinkhorn iteration}
\ENDIF
\ENDWHILE
 \RETURN $\gamma=\diag(a)\gamma^0\diag(b)$
\end{algorithmic}
\end{algorithm}

\begin{algorithm}
\caption{Over-relaxed Sinkhorn algorithm 2}
\label{SOR}
\begin{algorithmic}
\REQUIRE $\mu^1\in \IR^{n_1}$, $\mu^2\in \IR^{n_2}$, $c\in \IR^{n_1\times n_2}_+$
\STATE Set $a=\mathbf{1}_{n_1}$, $b=\mathbf{1}_{n_2}$, $\gamma^0=e^{-c/\epsilon}$, $\omega\in[1;2[$ and $\eta>0$
\WHILE {$||a\otimes \gamma^0b-  \mu_1||>\eta$}
\STATE $\tilde a=\mu_1\oslash (\gamma^0 b)$, 
\IF{ $\langle \mu^1,\varphi_\omega(a\oslash\tilde a) \rangle>0$}
\STATE  $a=a^{1-\omega}\otimes \tilde a^\omega$\COMMENT{Over-relaxed Sinkhorn iteration}
\ELSE
\STATE $a=\tilde a$,\COMMENT{Sinkhorn iteration}
\ENDIF
\STATE $\tilde b=\mu_2\oslash (^t\gamma^0  a)$
\IF{ $\langle \mu^2,\varphi_\omega(b\oslash\tilde b) \rangle>0$}
\STATE  $b=b^{1-\omega}\otimes \tilde b^\omega$\COMMENT{Over-relaxed Sinkhorn iteration}
\ELSE
\STATE $b=\tilde b$,\COMMENT{Sinkhorn iteration}
\ENDIF
\ENDWHILE
 \RETURN $\gamma=\diag(a)\gamma^0\diag(b)$
\end{algorithmic}
\end{algorithm}

>>>>

%{\color{red} 
%	TODO Construire les bonnes fonctions $\theta_k(\gamma)$.}
To construct the functions $\theta_k$ needed by the theorem, we need the following lemma.

@@

%In practice, we observed that the algorithm introduced in Theorem \ref{thm:algo} can be simplified and it is sufficient to check $F(P^\omega_{\Ccal_2}(P^\omega_{\Ccal_1}(\gamma^l)))-F(\gamma^l)>0$ for performing the over-relaxed iterations with parameter $\omega$.
%The cost of this test is linear with respect to the data dimension (i.e. $o(n_1+n_2)$).

\renewcommand{\algorithmiccomment}[1]{\hfill\bgroup(#1)\egroup}
%\begin{algorithm}
%\caption{Over-relaxed Sinkhorn algorithm}
%\label{SOR}
%\begin{algorithmic}
%\REQUIRE $\mu^1\in \IR^{n_1}$, $\mu^2\in \IR^{n_2}$, $c\in \IR^{n_1\times n_2}_+$
%\STATE Set $a=\mathbf{1}_{n_1}$, $b=\mathbf{1}_{n_2}$, $\gamma^0=e^{-c/\epsilon}$, $\omega\in[1;2[$ and $\eta>0$
%\WHILE {$||a\otimes \gamma^0b-  \mu_1||>\eta$}
%\STATE $\tilde a=\mu_1\oslash (\gamma^0 b)$,  $a_\omega=a^{1-\omega}\otimes \tilde a^\omega$
%\STATE $\tilde b=\mu_2\oslash (^t\gamma^0  a_\omega)$
%\IF{ $\langle \mu^1,\varphi_\omega(a\oslash\tilde a) \rangle +\langle \mu^2,\varphi_\omega(b\oslash\tilde b)\rangle>0$}
%\STATE  $a=a_\omega$, $b=b^{1-\omega}\otimes \tilde b^\omega$\COMMENT{Over-relaxed Sinkhorn iteration}
%\ELSE
%\STATE $a=\tilde a$, $b=\mu_2\oslash (^t\gamma^0 a)$  \COMMENT{Sinkhorn iteration}
%\ENDIF
%\ENDWHILE
% \RETURN $\gamma=\diag(a)\gamma^0\diag(b)$
%\end{algorithmic}
%\end{algorithm}
In order to summarize the whole algorithm, we first reformulate the functions $\theta_k(\gamma)$ defined in \eqref{eq:theta_k} as $\theta_k(\gamma)=\Theta(\frac{A_k \gamma}{\mu^k})$ where 
$ \Theta(u) = \min(\max(1,\Theta^*(u)-\delta),\theta_0)$ and  $\Theta^*(u)=\sup \left\{\omega \in [1;2]  \mid \varphi_\omega\left(\min u \right) \ge 0 \right\}$.

\begin{algorithm}
\caption{Over-relaxed Sinkhorn algorithm }
\label{SOR}
\begin{algorithmic}
\REQUIRE $\mu^1\in \IR^{n_1}$, $\mu^2\in \IR^{n_2}$, $c\in \IR^{n_1\times n_2}_+$
\STATE Set $a=\mathbf{1}_{n_1}$, $b=\mathbf{1}_{n_2}$, $\gamma^0=e^{-c/\epsilon}$, $\theta_0\in[1;2)$ and $\eta>0$
\WHILE {$||a\otimes \gamma^0b-  \mu_1||>\eta$}
\STATE $\tilde a=\mu_1\oslash (\gamma^0 b)$, 
\STATE $\omega=\Theta(a\oslash\tilde a)$
%\IF{ $\langle \mu^1,\varphi_\omega(a\oslash\tilde a) \rangle>0$}
\STATE  $a=a^{1-\omega}\otimes \tilde a^\omega$%\COMMENT{Over-relaxed Sinkhorn iteration}
%\ELSE
%\STATE $a=\tilde a$,\COMMENT{Sinkhorn iteration}
%\ENDIF
\STATE $\tilde b=\mu_2\oslash (^t\gamma^0  a)$
\STATE $\omega=\Theta(b\oslash\tilde b)$
%\IF{ $\langle \mu^2,\varphi_\omega(b\oslash\tilde b) \rangle>0$}
\STATE  $b=b^{1-\omega}\otimes \tilde b^\omega$%\COMMENT{Over-relaxed Sinkhorn iteration}
%\ELSE
%\STATE $b=\tilde b$,\COMMENT{Sinkhorn iteration}
%\ENDIF
\ENDWHILE
 \RETURN $\gamma=\diag(a)\gamma^0\diag(b)$
\end{algorithmic}
\end{algorithm}
